---
title: "Chapter 8 Notes"
author: "Tim"
date: "10/10/2017"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, 
                      results='show', cache=TRUE, autodep=TRUE)
```

## 8.3 Models with Stan
```{r}
library(rethinking)
data(rugged)
d <- rugged
d$log_gdp <- log(d$rgdppc_2000)
dd <- d[complete.cases(d$rgdppc_2000),]
```

Stan requires 'trimmed' variable sets that only contain the necessary data:

```{r}
dd.trim <- dd[,c("log_gdp","rugged","cont_africa")]
```
And we can fit the model with Stan:
```{r, results='hide'}
m8.1stan <- map2stan(
  alist(
    log_gdp <- dnorm(mu, sigma),
    mu <- a + bR*rugged + bA*cont_africa + bAR*rugged*cont_africa,
    a ~ dnorm(0,100),
    bR ~ dnorm(0,10),
    bA ~ dnorm(0,10),
    bAR ~ dnorm(0,10),
    sigma ~ dcauchy(0,2)
  ),
  data=dd.trim
)
```
```{r}
precis(m8.1stan)
```
We can sample:
```{r}
post <- extract.samples(m8.1stan)
str(post)
```
And
```{r}
pairs(m8.1stan)
```
We also have these utility functions:
```{r}
WAIC(m8.1stan)
show(m8.1stan)
```
Trace plot:
```{r}
plot(m8.1stan)
```
### Debugging
```{r, results='hide'}
y <- c(-1,1)
m8.2 <- map2stan(
  alist(
    y ~ dnorm(mu,sigma),
    mu <- alpha
  ),
  data=list(y=y), start=list(alpha=0, sigma=1),
  chains=2, iter=4000, warmup=1000)
```
```{r}
precis(m8.2)
```
```{r}
plot(m8.2)
```


We get a completely wrong solution. We just need to try some weakly informative priors:
```{r, results='hide'}
m8.3 <- map2stan(
  alist(
    y ~ dnorm(mu,sigma),
    mu <- alpha,
    alpha ~ dnorm(1, 10),
    sigma ~ dcauchy(0,1)
  ),
  data=list(y=y), start=list(alpha=0, sigma=1),
  chains=2, iter=4000, warmup=1000)
```
```{r}
precis(m8.3)
```
```{r}
plot(m8.3)
```

### Non-identifiable parameters
We see what happens when we try to fit highly correlated predictors and non-identifiable parameters:

```{r}
y <- rnorm(100, mean=0, sd=1)
```
```{r, results='hide'}
m8.4 <- map2stan(
  alist(
    y ~ dnorm(mu, sigma),
    mu <- a1 + a2,
    sigma ~ dcauchy(0,1)  
  ),
  data=list(y=y), start=list(a1=0, a2=0, sigma=1),
  chains=2, iter=4000, warmup=1000)
```
```{r}
precis(m8.4)
```
WE see terrible n_eff and Rhat values. 

```{r}
plot(m8.4)
```

The parameters are not stationary.

Issues like this can be solved by adding weakly informative priors. This is good advice in general; it helps to fit models and make better decision. If for example, 30 million isn't a value model, don't use flat priors.
```{r, results='hide'}
m8.5 <- map2stan(
  alist(
    y ~ dnorm(mu, sigma),
    mu <- a1 + a2,
    a1 ~ dnorm(0,10),
    a2 ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)  
  ),
  data=list(y=y), start=list(a1=0, a2=0, sigma=1),
  chains=2, iter=4000, warmup=1000)
```
```{r}
precis(m8.5)
```


